Technique for \earlywarning\ detection of compact binary coalescence and its implication for multi-messenger astronomy

### 2011-06-03 implementation.tex
Renamed analysis.tex to implementation.tex
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@893 ccdd05d5-56c1-461a-bf98-85f6876523b4
Edited section 3.1
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@897 ccdd05d5-56c1-461a-bf98-85f6876523b4

In this section we describe an implementation of the \lloid\ method described
in section \ref{sec:method} suitable for rapid gravitational-wave
searches for compact binary coalescence.  The \lloid\ method requires several
computations that can be completed before the analysis is underway.  Thus
we divide the procedure into two stages 1) an off-line planning stage and 2) an
online, low-latency filtering stage.  The off-line stage can be done before the
analysis is started and updated asynchronously, whereas the online stage must
keep up with the detector output and produce search results as rapidly as
possible.  In the next two subsections we describe what these stages entail.

@@

The choice of filter waveforms and singular value decomposition can be done in
advance and will be valid as long as the detector noise spectrum remains
roughly constant.  New filter waveforms can be computed asynchronously using
updated spectrum estimates as they are available.

\begin{figure}[htbp]
	\label{fig:tmpltbank}
	\begin{center}
		\includegraphics[scale=0.75]{figures/tmpltbank.png}
		\caption{Placement of template parameters used in this paper.  The template bank consists of 98544 templates with component masses $m_1$, $m_2$, between 1~and~3~$M_\odot$.  We shall design a filter bank to search for a small subset of these template parameters with chirp masses $\mathcal M$ between 1.1955~and~1.2045~$M_\odot$.}
	\end{center}
\end{figure}

\begin{table}
\begin{indented}
\caption{Filter design for these 657 templates.  From left to right, this table shows the sample rate, time interval, number of samples, and number of orthogonal templates for each time slice.  The \textsc{svd} tolerance is varied from $\left(1-10^{-1}\right)$ to $\left(1-10^{-6}\right)$}
\item[]\begin{tabular}{rr@{,}lc*{6}{r}}
\br
\multicolumn{4}{c}{} &\centre{6}{$\log_{10}$ (1 - \textsc{svd} tolerance)} \\
\ns
\multicolumn{4}{c}{} &\crule{6} \\
$f^s$ (Hz) & $(t^{s+1}$&$t^s]$ (s) & $N^s$ & -1 & -2 & -3 & -4 & -5 & -6 \\
\mr
4096 & (0.5&0] & 2048 & 1 & 4 & 6 & 8 & 10 & 14 \\
512 & (4.5&0.5] & 2048 & 2 & 6 & 8 & 10 & 12 & 16 \\
256 & (12.5&4.5] & 2048 & 2 & 6 & 8 & 10 & 12 & 15 \\
128 & (76.5&12.5] & 8192 & 6 & 20 & 25 & 28 & 30 & 32 \\
64 & (140.5&76.5] & 4096 & 1 & 8 & 15 & 18 & 20 & 22 \\
64 & (268.5&140.5] & 8192 & 1 & 7 & 21 & 25 & 28 & 30 \\
64 & (396.5&268.5] & 8192 & 1 & 1 & 15 & 20 & 23 & 25 \\
32 & (460.5&396.5] & 2048 & 1 & 1 & 3 & 9 & 12 & 14 \\
32 & (588.5&460.5] & 4096 & 1 & 1 & 7 & 16 & 18 & 21 \\
32 & (844.5&588.5] & 8192 & 1 & 1 & 8 & 26 & 30 & 33 \\
32 & (1100.5&844.5] & 8192 & 1 & 1 & 1 & 12 & 20 & 23 \\
\end{tabular}
\end{indented}
\end{table}

The planning stage begins with choosing templates that cover the space of mass
parameters with a hexagonal grid~\cite{PhysRevD.76.102004} in order to satisfy
a minimum match criterion.  This assures a user-specified, maximum loss in \SNR\
for signals that fall in-between the chosen templates.  Typically the minimum
match is 97\% corresponding to a maximum mismatch of 3\%.  Next, the templates
are subdivided into groups of neighbors called ``sub-banks'' that are
appropriately sized so that each bank can be efficiently handled by a single
computer.  The neighbors are chosen to have comparable chirp mass, which produces
sub-banks with very similar waveforms.  Dividing the mass space into smaller
sub-banks reduces the computational cost of the singular value decomposition
and is the approach considered in~\cite{Cannon:2010p10398}.  Using our
understanding of the time-frequency evolution of the templates, we choose time
slice boundaries as in \eqref{FIXME} such that all of the templates within a
sub-bank are sub-critically sampled at progressively lower sample rates.  Next,
the templates within the sub-bank are realized as \fir\ filter
coefficients.  For each time slice, the templates are down-sampled to the
appropriate sample rate.  Finally, the \SVD\ is applied to each time
slice in the sub-bank in order to produce a set of orthogonal \fir\
filters and a reconstruction matrix that maps them back to the original
templates as described in \eqref{FIXME}.  The down-sampled orthogonal
\fir\ filter coefficients, the reconstruction matrix, and the time slice
boundaries are all saved to disk.

>>>>

In this section we describe an implementation of the \lloid\ method described
in section \ref{sec:method} suitable for rapid gravitational-wave
searches for compact binary coalescence.  The \lloid\ method requires several
computations that can be completed before the analysis is underway.  Thus
we divide the procedure into two stages 1) an offline planning stage and 2) an
online, low-latency filtering stage.  The offline stage can be done before the
analysis is started and updated asynchronously, whereas the online stage must
keep up with the detector output and produce search results as rapidly as
possible.  In the next two subsections we describe what these stages entail.

@@

The choice of templates and the \SVD\ can be done in
advance and will be valid as long as the detector noise spectrum remains
roughly constant%
\editorial{I'd rather not go into this.}%
.  New templates and the \SVD\ can be computed asynchronously using
updated spectrum estimates as they are available.

\begin{figure}[htbp]
	\label{fig:tmpltbank}
	\begin{center}
		\includegraphics{figures/tmpltbank.png}
		\caption{Placement of template parameters used in this paper.  The template bank consists of 98544 templates with component masses $m_1$, $m_2$, between 1~and~3~$M_\odot$.  We shall design a filter bank to search for a small subset of these template parameters with chirp masses $\mathcal M$ between 1.1955~and~1.2045~$M_\odot$.}
	\end{center}
\end{figure}

\begin{table}
\begin{indented}
\caption{Filter design for these 657 templates.  From left to right, this table shows the sample rate, time interval, number of samples, and number of orthogonal templates for each time slice.  The \textsc{svd} tolerance is varied from $\left(1-10^{-1}\right)$ to $\left(1-10^{-6}\right)$}
\item[]\begin{tabular}{rr@{,}lc*{6}{r}}
\br
\multicolumn{4}{c}{} &\centre{6}{$\log_{10}$ (1 - \textsc{svd} tolerance)} \\
\ns
\multicolumn{4}{c}{} &\crule{6} \\
$f^s$ (Hz) & $(t^{s+1}$&$t^s]$ (s) & $N^s$ & -1 & -2 & -3 & -4 & -5 & -6 \\
\mr
4096 & (0.5&0] & 2048 & 1 & 4 & 6 & 8 & 10 & 14 \\
512 & (4.5&0.5] & 2048 & 2 & 6 & 8 & 10 & 12 & 16 \\
256 & (12.5&4.5] & 2048 & 2 & 6 & 8 & 10 & 12 & 15 \\
128 & (76.5&12.5] & 8192 & 6 & 20 & 25 & 28 & 30 & 32 \\
64 & (140.5&76.5] & 4096 & 1 & 8 & 15 & 18 & 20 & 22 \\
64 & (268.5&140.5] & 8192 & 1 & 7 & 21 & 25 & 28 & 30 \\
64 & (396.5&268.5] & 8192 & 1 & 1 & 15 & 20 & 23 & 25 \\
32 & (460.5&396.5] & 2048 & 1 & 1 & 3 & 9 & 12 & 14 \\
32 & (588.5&460.5] & 4096 & 1 & 1 & 7 & 16 & 18 & 21 \\
32 & (844.5&588.5] & 8192 & 1 & 1 & 8 & 26 & 30 & 33 \\
32 & (1100.5&844.5] & 8192 & 1 & 1 & 1 & 12 & 20 & 23 \\
\br
\end{tabular}
\end{indented}
\end{table}

The planning stage begins with choosing templates that cover the space of source
parameters with a hexagonal grid~\cite{PhysRevD.76.102004} in order to satisfy
a minimum match criterion.  This assures a prescribed maximum loss in \SNR\
for signals that fall between the chosen templates.  Typically the minimum
match is 0.97 corresponding to a maximum mismatch of 0.03.  Next, the templates
are subdivided into groups of neighbors called \emph{sub-banks} that are
appropriately sized so that each bank can be efficiently handled by a single
computer.  The neighbors are chosen to have comparable chirp mass, which produces
sub-banks with similar time-frequency evolution.  Dividing the source parameter space
into smaller sub-banks reduces the computational cost of the singular value decomposition
and is the approach considered in~\cite{Cannon:2010p10398}.  We choose time
slice boundaries as in equation~\eqref{eq:time-sliced-templates} such that all of the templates within a
sub-bank are sub-critically sampled at progressively lower sample rates.  Next,
the templates within the sub-bank are realized as \fir\ filter
coefficients.  For each time slice, the templates are down-sampled to the
appropriate sample rate.  Finally, the \SVD\ is applied to each time
slice in the sub-bank in order to produce a set of orthogonal \fir\
filters and a reconstruction matrix that maps them back to the original
templates as described in equation~\eqref{eq:svddecomp}.  The down-sampled orthogonal
\fir\ filter coefficients, the reconstruction matrix, and the time slice
boundaries are all saved to disk.
